---
title: "linear_regression.R"
author: "Alrik Schörling"
date: '2022-11-14'
output: pdf_document
---

Set up
```{r setup}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE)

knitr::opts_knit$set(root.dir = "/Users/al7743sc/Dropbox (Human Neural Develop)/Alrik Kirkeby PhD/Miscellaneous/Students/Erno 2022")
```

Install and load packages
```{r}
#install.packages("packman") #run if you don't have
#pacman::p_install(dplyr, tidyverse, ggplot2, rstatix) #run if you don't have
#pacman::p_load(dplyr, tidyverse, ggplot2, rstatix) #load packages 
```

Create useful themes
```{r}
basic_theme <-  theme_bw() + 
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.ticks = element_line(size = 0.5),
        axis.ticks.length = unit(.2, "cm"), 
        axis.text.x = element_text(color = "black", family = "sans", size = 12),
        axis.text.y = element_text(color = "black", family = "sans", size = 12),
        axis.title = element_text(size = 12), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        strip.text.x = element_text(size = 12),
        strip.text.y = element_text(size = 12), 
        strip.background = element_rect(color = NA, fill = NA), 
        plot.title = element_text(color = "black", family = "sans",  size = 12))

theme_1 <- basic_theme + theme(axis.line = element_line(size = 0.5))

theme_2 <- basic_theme + 
  theme(axis.line = element_line(size = 0.5), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

theme_3 <- basic_theme + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 1), 
        panel.border = element_rect(size = 1, fill = NA))

theme_4 <- basic_theme + 
  theme(panel.border = element_rect(size = 1, fill = NA))

theme_5 <- basic_theme + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), 
        panel.border = element_rect(size = 1, fill = NA))
```

Load data an convert to long format
```{r}
erno_corr_long <- read.csv("yz004nr if p1.csv", sep = ",") |>
  select(-graft_no) |> 
  mutate(image_id = seq(1:n())) |>
  pivot_longer(LMX:harmony_th, values_to = "count", names_to = "marker") |>
  mutate(marker = tolower(marker)) |> 
  mutate(count_by = case_when(marker == "lmx" | marker == "th" ~ "manual",
                              marker == "sc_lmx" | marker == "sc_th" ~ "simplicity", 
                              marker == "harmony_lmx" | marker == "harmony_th" ~ "harmony")) |>
  mutate(marker = gsub("sc_", "", marker), 
         marker = gsub("harmony_", "", marker)) |>
  mutate(marker = factor(marker)) |> 
  mutate(count_by = factor(count_by)) |> 
  drop_na() |>
  arrange(count)

#inspect the data
head(erno_corr) 
glimpse(erno_corr)
```

Inspect normality (requirment for parametric, linear correlation)
```{r}
erno_corr_long |> 
  group_by(count_by) |>
  shapiro_test(count)

ggplot(erno_corr_long, aes(x = count)) + 
  geom_histogram(aes(y =..density..), fill = "gray90") +
  geom_density(aes(fill = count_by), size = 2, alpha = 0.5) + 
  xlab("") + 
  scale_x_continuous(limits = c(-200, 1700)) +
  theme_1
```

Some have shapiro-wilk p<0.05 but the distributions are relatively normal. Proceed with parametric linear regression. Correlation is sometimes easier done in a wide data format, where the column corresponds to the previous catagorical data. Convert to wide format 
```{r}
erno_corr_wide <- erno_corr |> 
  pivot_wider(values_from = "count", names_from = "count_by")

head(erno_corr_wide) #inspect the data 
```


```{r}
#corr manual ~ harmony 
lm_h <- lm(erno_corr_wide$manual ~ erno_corr_wide$harmony, erno_corr_wide)
glance(lm_h) #

#corr manual ~ simplicity
lm_s <- lm(erno_corr_wide$manual ~ erno_corr_wide$simplicity, erno_corr_wide)
glance(lm_s)
```
Simplicity performs quite poorly based upon the r-squared. Let's plot the data to see what's up

```{r}
corr_plt <- function(data, x_var, y_var) {
  ggplot(data, aes(x = {{x_var}}, y = {{y_var}})) + 
    geom_point(shape = 21, size = 2) + 
    geom_smooth(method = "lm", se = F) + 
    facet_wrap(~marker) + 
    theme_1
  }

corr_plt(erno_corr_wide, manual, harmony)
corr_plt(erno_corr_wide, manual, simplicity)
```

It seems simplicity underestimates the number of cells when bigger than >500 in the manual count. Let's filter these out and see what happens. The Harmony has too few points to correlate imo

```{r}
erno_corr_wide_sel <- erno_corr_wide |> 
  filter(manual < 400)
```

```{r}
corr_plt(erno_corr_wide_sel, manual, harmony)
corr_plt(erno_corr_wide_sel, manual, simplicity)
```

Looks quite a lot better! Let's look at the stats
```{r}
lm_s2 <- lm(erno_corr_wide_sel$manual ~ erno_corr_wide_sel$simplicity, erno_corr_wide_sel)
glance(lm_s2)
```

Let's plot SIMPLIcity without facet_wrap
```{r}
corr_plt_alt <- function(data, x_var, y_var) {
  ggplot(data, aes(x = {{x_var}}, y = {{y_var}})) + 
    geom_smooth(method = "lm") + 
    geom_point(aes(fill = marker), shape = 21, size = 2) + 
    geom_hline(yintercept = c(100, 200, 300), linetype = 2, size = 0.2) +
    geom_vline(xintercept = c(100, 200, 300), linetype = 2, size = 0.2) +
    scale_y_continuous(limits = c(0, 400), exp = c(0,0)) + 
    scale_x_continuous(limits = c(0, 400), exp = c(0,0)) + 
    theme_1
  }

plt <- corr_plt_alt(erno_corr_wide_sel, manual, simplicity)

plt
ggsave("erno_corr_wide_sel.pdf", plt, height = 4, width = 5)
```
